# Analysis on Weight Lifting Exercise Dataset

## Executive Summary

We analyzed the data to see if it is possible to predict which activity was performed at specific point in time.

## Load the training set and transform it to fit our needs

The data, Weight Lifting Exercise Dataset, is from http://groupware.les.inf.puc-rio.br/har.

```{r, cache=TRUE}
url <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'

# Do not transform string values into factors:
data <- read.csv(url, na.strings=c("", "NA"), stringsAsFactors=FALSE)
```

# Data Cleaning

See Appendix for data description.

Check that variable classe is a factor:

```{r}
class(data$classe)
```

Transform it into a factor:

```{r}
data$classe <- as.factor(data$classe)
class(data$classe)
```

```{r}
# save the original ncol value for later checkup
originalNCol <- ncol(data)
originalNCol
```

Remove unnecessary columns
```{r}
data <- data[, -c(1,2,3,4,5,6,7)]
```


Remove columns filled with NA's

```{r}
# we need a function that counts the NA's
howmanyNAs <- function(colName) { length(which(is.na(data[, colName]))) }

# then, let's count the NA's and put the result into a table
nainfo <- sapply(names(data), howmanyNAs)

# to keep things clear, let's make a list of column names that are to be removed:
removeThese <- names(nainfo[nainfo > 19000])

# finally, we are ready to remove the columns
data <- data[, -which(names(data) %in% removeThese)]

# Check that numbers match
originalNCol
ncol(data)
originalNCol - ncol(data)
length(removeThese)

```

# Random Forest Model

```{r, cache=TRUE}
set.seed(100)
library(randomForest)
rfModel <- randomForest(classe ~., data=data)
print(rfModel)
```

The OOB estimate of error rate is 0.29%. 

## Variable Importance

```{r, cache=TRUE}
varImpPlot(rfModel)
```



# Load Testing Data

```{r, cache=TRUE}
url_testData <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

# Do not transform string values into factors:
testData <- read.csv(url_testData, na.strings=c("", "NA"), stringsAsFactors=FALSE)
```

# Predict with the Model

```{r, cache=FALSE}
answers <- predict(rfModel, testData)
```

# Create submission files

```{r, cache=TRUE}

pml_write_files = function(x)
{
  n = length(x)
  for(i in 1:n)
  {
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

``` 
Then submit them manually. Yes, manually.

# Appendix

## Data description

```{r}
str(data)
```
